---
title: "p8105_hw2_th3147"
author: "Te-Hsuan Huang"
date: "2025-09-27"
output: github_document
---

# Problem 1

1. First dataset

```{r}
library(tidyverse)

polsmonth_df = read_csv(file = "./fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(mon, into = c("year", "month", "day"), sep = "-") |> 
 mutate( 
   year = as.numeric(year),
    month = case_match(
      month,
      "01" ~ "jan",
      "02" ~ "feb",
      "03" ~ "mar",
      "04" ~ "apr",
      "05" ~ "may",
      "06" ~ "jun",
      "07" ~ "jul",
      "08" ~ "aug",
      "09" ~ "sep",
      "10" ~ "oct",
      "11" ~ "nov",
      "12" ~ "dec"
    ), 
    
    president = case_match(
      prez_dem,
      1 ~ "dem",
      0 ~ "gop")) |> 
    select(-prez_dem, -prez_gop, -day)

polsmonth_df
  
```

* make sure the lower case in month.

2. second dataset

```{r}
snp_df = read_csv(file = "./fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |> 
  separate(date, into = c("month", "day", "year"), sep = "/") |>
  mutate(
    year = as.integer(paste0("20", year)),
    year = as.numeric(year),
    month = case_match(
      month,
      "1" ~ "jan",
      "2" ~ "feb",
      "3" ~ "mar",
      "4" ~ "apr",
      "5" ~ "may",
      "6" ~ "jun",
      "7" ~ "jul",
      "8" ~ "aug",
      "9" ~ "sep",
      "10" ~ "oct",
      "11" ~ "nov",
      "12" ~ "dec"
    )) |> 
      select(-day) |>
      arrange (year, month) |> 
      relocate(year, month)

snp_df
```

* paste0: is a function in R used to concatenate (join together) character strings.

* make sure the lower case in month.


3. Third dataset

```{r}
unemploy_df = read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment_rate") |> 
  relocate(year, month) |>
  arrange(year, month)

unemploy_df
```

4. Join the dataset

```{r}
final_merged_df = 
   polsmonth_df |>
  left_join(snp_df, by = c("year", "month")) |>
  left_join(unemploy_df, by = c("year", "month"))

final_merged_df
```

* Keep in mind: 

1. the features of year and month in each dataset should be same.

2. the alphabet should be exactly same in month.

* The resulting dataset is a time-series data frame created by merging three original sources: pols-month.csv, which contains U.S. political control data (including the number of Democratic and Republican governors and the party of the sitting President); snp.csv, which tracks the S&P 500 closing price; and unemployment.csv, which records the national unemployment rate. After joining the three files on the common keys of year and month, the final merged dataset contains 822 rows and 11 columns, covering the time span from 1947 to 2015. Key variables include the date identifiers (year, month), political metrics (gov_gop, rep_dem, president), the financial index (close), and the economic indicator (unemployment_rate).


# Problem 2

1. First dataset

```{r}
mrtrash_df = 
  readxl::read_excel("./problem 2/202409 Trash Wheel Collection Data.xlsx"
                     , sheet="Mr. Trash Wheel"
                     , range = "A2:N653") |> 
  janitor::clean_names() |> 
   mutate(
    sports_balls = as.integer(round(as.numeric(`sports_balls`))),
    wheel_name = "Mr. Trash Wheel",
    year=as.numeric(year)
  ) |> 
  select(-dumpster) |> 
  relocate(year, month, wheel_name) |> 
  arrange(date)

mrtrash_df
```

2. Second dataset

```{r}
prtrash_df = 
  readxl::read_excel("./problem 2/202409 Trash Wheel Collection Data.xlsx"
                     , sheet="Professor Trash Wheel"
                     , range = "A2:M120") |> 
  janitor::clean_names() |> 
  mutate(
    wheel_name = "Professor Trash Wheel") |> 
  select(-dumpster) |> 
  relocate(year, month, wheel_name) |> 
  arrange(date)

prtrash_df
```

3. Third dataset

```{r}
gwtrash_df = 
  readxl::read_excel("./problem 2/202409 Trash Wheel Collection Data.xlsx"
                     , sheet="Gwynnda Trash Wheel"
                     , range = "A2:L265") |> 
  janitor::clean_names() |> 
  mutate(
    wheel_name = "Gwynnda Trash Wheel") |> 
  select(-dumpster) |> 
  relocate(year, month, wheel_name) |> 
  arrange(date)
  
gwtrash_df
```

4. Merge the dataset

```{r}
final_merged2_df = 
  bind_rows(mrtrash_df, prtrash_df, gwtrash_df) |>
  janitor::clean_names()

final_merged2_df
```

